
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```


# Data Description:
The data used for this blog post was obtained from Our World in Data in the form of a .csv. The data contains information about new covid cases, vaccination numbers, icu patient numbers and death numbers, for each day between 13/03/2020 and 27/08/2021 for several hundred countries. The data for Kazakhstan was extracted from this csv file and saved locally.

Citation:

  Hannah Ritchie, Edouard Mathieu, Lucas Rodés-Guirao, Cameron Appel, Charlie Giattino, Esteban Ortiz-Ospina, Joe Hasell, Bobbie Macdonald, Diana Beltekian and Max Roser (2020) - "Coronavirus Pandemic (COVID-19)". Published online at OurWorldInData.org. Retrieved from: 'https://ourworldindata.org/coronavirus' [Online Resource]

# Kazakhstan: Summers are for infection

Kazakhstan, like the rest of the world, has been deeply effected by the COVID-19 pandemic. Their infection rates tell a story of 2020 devastation, followed by a 2021 catastrophe. 

Aside from the overall surge of cases in 2021, it is worth noting that the deadlier months tend to be in the Summer - with the most number of cases cropping up in July and August 2021, and with many deaths recorded in both August 2020 and August 2021.

This can be shown by playing around with the arrows at the top of the New_Cases and New_Deaths columns in the first table.


```{r}
library(tidyverse)
library(dplyr)
library(DT)
library(tidyverse)
library(lubridate)
library(formattable)
library(gt)

raw <- read.csv("kazakhstan_data.csv")
raw_ <- raw %>% select(
  date, total_cases, new_cases, total_deaths, new_deaths, reproduction_rate, new_tests, total_tests, total_vaccinations,people_vaccinated, people_fully_vaccinated, total_boosters, new_vaccinations,population) %>% 
  mutate(percentage_vaxed = people_fully_vaccinated/population, 
         percentage_dead = total_deaths/population, 
         percentage_cases = total_cases/population, 
         month= months(dmy(date)),
         monthref = month(dmy(date)),
         year = year(dmy(date))
  )
      
table1 <- raw_ %>% 
  select(year, monthref, month, total_cases, new_cases, new_deaths) %>% 
  group_by(year, monthref, month) %>% 
  summarise(New_Cases = sum(new_cases), New_Deaths=sum(new_deaths)) %>%
  datatable(options=list(order=list(1, "asc"),list(list(4), list(5) )))



vax <- raw_[406:533,]

vax_<- vax %>% 
  select(date, people_vaccinated, population, year, month) %>% mutate(percentage=round(people_vaccinated/population,digits=4)) %>%
  fill(people_vaccinated, .direction="up") %>% 
  fill(percentage, .direction="up") %>% 
  group_by(year, month) %>% summarise(Total.Vaccinations= max(people_vaccinated), Percentage.Vaccinated=max(percentage)) %>% 
  arrange(Percentage.Vaccinated)


table2 <- vax_[, 2:4] %>% 
  gt() %>% 
  tab_header(title="2021 Kazakhstan Cumulative Vaccinations", 
             subtitle = "Vaccination numbers since the first rollout of vaccinations in April") %>%
  tab_source_note("Source: Our World in Data 2021")

```




# Cases and Deaths

### Total Cases and Deaths per month
#### order by time, cases or deaths as you like!


```{r}
table1
```


# Vaccinations
In combating the virus, Kazakhstan started their vaccination rollout in April 2021.

Vaccination rates have brought them from 7% vaccinated to 34% vaccinated in 5 months, as seen by the final column in the table below.

```{r}
table2
```


